\documentclass[letterpaper]{article} % Feel free to change this

\usepackage{amsmath}

\begin{document}

\title{ECE 350: Digital Systems Lab 5}
\author{Walker Willetts and Joseph Dechicchis} % Change this to your name(s)
\date{\today} % Change this to the date you are submitting
\maketitle

\section*{Duke Community Standard}

By submitting this \LaTeX{} document, I affirm that
\begin{enumerate}
    \item I understand that each \texttt{git} commit I create in this repository is a submission
    \item I affirm that each submission complies with the Duke Community Standard and the guidelines set forth for this assignment
    \item I further acknowledge that any content not included in this commit under the version control system cannot be considered as a part of my submission.
    \item Finally, I understand that a submission is considered submitted when it has been pushed to the server.
\end{enumerate}
\newpage

\section{Introduction}
    The overall goal of the lab was to familiarize ourselves with the FPGA hardware, and to demonstrate the workflow for uploading and operating the hardware. The FPGA hardware was first explored through an activity involving the display of a video to a monitor using the VGA connector on the FPGA board. During this component of the lab, the process for utilizing memory with the board was examined as well as the setup of inputs such as buttons to the board. In the second section of the lab, we explored how to handle the conversion of between different encodings of I/O data sent to the board using programmed logic in verilog.

\section{Questions}
    \subsection{Question 1: Interface vs. Controller Module}
        An interface module offers an interface for abstracting away interactions with a more complicated piece of hardware such as a VGA output or RAM. A controller module, on the other hand, defines the actual piece of complex hardware described above with all of its complexities. Interface modules will implement a controller module that represents the complex hardware that the interface abstracts for other modules in order to make manipulations such as screen colors via VGA far simpler. 

        In general, a controller module is first implemented in order provide a manner for recieving output from and sending input to some piece of hardware. An interface module is then built that performs the complex manipulations on the hardware needed in order to obtain desired behavior. This interface is then used in other modules as a simple means of performing the hardware manipulation.
    
    \subsection{Question 2: MVC Relationships} 

    \subsection{Question 3: Memory-Mapped IO Operation}
        Memory mapped IO works by mapping I/O devices to a memory address within the address space of a functional unit, such as a CPU. The I/O device can then be accessed as if it were just data in RAM allowing for both writing and reading functionality. The access to these addresses by the functional unit are monitored by the I/O devices, which then respond to the access appropriately.
    
    \subsection{Question 4: Memory-Mapped IO and Load/Store Instructions}
        Memory-mapped IO relates to the load/store assembly instructions because it provides an abstraction that allows I/O to devices to be replaced by either calling a load word instruction to the address used to represent the device, which triggers the IO device to provide data to read, or by calling a store word instruction to the address used to represent the device, which triggers the IO device to read provided data. 
        
        640
    \subsection{Question 5: VGA Clock}
        Assuming a 640 by 640 VGA display with a 60 Hz refresh rate:
        \begin{align*}
            VGA_{clock} &= pixels * refresh_rate \\
            VGA_{clock} &= (640 * 640 pixels) * (60 Hz) \\
            VGA_{clock} &= 24576000 Hz = 24.576 MHz
        \end{align*}
        However, this calculation does assume that there is no included overscan pixels for the monitor. 

    \subsection{Question 6: Displaying Image MIF}
        In order to display a moving image instead of the square, there would need to be two things. First, some sort of anchor point would be needed in the image, which would be used in generating pixel values similar to the top left corner of the square generated in lab. Second, there would need to be a mapping provided for the images pixels that give their coordinate relative to the anchor point, which could be provided using memory. Using both these, displaying the images would involve calculating the coordinates of a pixel on the screen by means of the saved anchor point, and then setting this pixels RGB value appropriately.
        
    \subsection{Question 7: Lab Improvements}
        For the most part, the lab seemed to be pretty well done, and it demonstrated a lot of useful manipulations for the FPGA board. The last section with the PS1 controller is really the only thing that seemed a little pointless due to its brevity.
\section{Lab Components and Decisions}
    \subsection{Generating Colors Using FPGA And VGA}
        This section of the lab was relatively straightforward as it only required coloring two halves of a monitor using a .MIF file. In order to accomplish this task, we first altered the img\_data.v file to use a custom .MIF file where the first 2000 or so memory addresses contained only zero values while the remaining addresses contained only the value one. Next, we altered the .MIF file containing the colors for the VGA controller to use the color orange for the for the pixels whose memory address in the first .MIF file were zero, and to use the color maroon for the pixels whose memory address were one. This is all that was needed in order to get the split colors on the screen.

    \subsection{Programming a Movable Square}
        In order to program a movable square, we first added a register to the VGA controller that contained the address in memory of the pixel that currently corresponded to the top left of the square. We then added a check before setting a pixel value at a given memory address that determines if the pixel value was within the bounds of the square, which we defined as being the 60 columns starting with the top left square pixel and going to the right as well as the 60 rows starting at the top left square pixel and going down. If the pixel was within these bounds, then we changed its value to be that of the color of the square. Next, we added a counter that allowed for smoother movement of the square by restricting how often updates to its top left corner position could occur. Finally, we added inputs corresponding to buttons on the board that were checked within the counter periodically to determine how to update the top left corner position of the square, which allowed for bidirectional movement. 

    \subsection{Examining the PS2 I/O Device}
        To convert the PS2 input values to the proper characters on the LCD, we first noted down the different hex values appearing on the segment displays that corresponded to the data sent by the PS2 to the FPGA board. We then added checks to determine if the values being sent by the PS2 corresponded to these hex values, and, if they did, we converted the PS2 data to be the proper ASCII encoding of the character. The ASCII encoding was then sent to the LCD that interpretted the value correctly, displaying the right character for the input key pressed on the PS2.
    \section{Insights Gained}
    Over the course of the lab, we gained a better understanding of the workflow for developing on an FPGA using Quartus, which includes the bottlenecks associated with the workflow. The slow pace at which code can actually be tested on the FPGA due to the compilation and upload process emphasized the added benefits of writing test benches because they greatly speed up development. The lab also provided a few example of how memory works with verilog and the FPGA, which was crucial to understanding how to incorporate memory with out CPU. 

\section{Challenges Faced}
    There were two main challenges that we faced when completing this lab. The first one was trying to determine an appropriate counter speed for the movement of the square. This was mainly a challenge due to the slow speed at which Quartus compiles code, which made narrowing down values take an unneccesary amount of time. The other challenge was trying to get input working for the FPGA, but this issue only happened because we accidentally were pressing buttons on the wrong board, leading us to think ours wasn't working when it was working.

\section{Conclusion}
        Overall, the lab provide a thorough examination of the FPGA board and how to accomplish various goals using the board. The lab will no doubt be an invaluable experience to reflect on when implementing the final processor project for the class.  

\end{document}
